---
title: "Preprocessing Data Analysis"
author: "Maria Calderon-Marrero"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(fig.path = "../figures/03_PreProcessing")

```

# Goals

1. Remove chloroplasts
2. Remove mitochondria
3. Remove ASVs from negative controls and also negative control samples.
4. Evaluate accuracy by looking at the Mock Community.
5. Check for reverse complements!
6. Remove samples without “enough” reads.

# Input

1. Raw phyloseq object

```
cp /workdir/in_class_data/raw_physeq.RData/workdir/mdc289/Biodiversity_Salinity_16S/data/02_AssignASVs
```

# Output

1. CLeaned up phyloseq object called 'data/03_PreProcessing/preprocessed_physeq.Rdata

#Set up the Environment

```{r load-packages-data}
set.seed(238428)

#Load packages
pacman::p_load(phyloseq, tidyverse, install = FALSE)

# Load data
load("data/02_AssignASVs/raw_physeq.RData")
#Take a look
raw_physeq
```

# CLean up data

## Chloroplasts & Mitochondria

```{r- chloromito}
noChloroMitos_physeq <-
raw_physeq %>%
  subset_taxa(Order != "Chloroplast" | is.na(Order)) %>% #248 cholorplasts
 subset_taxa(Family != "Mitochondria" | is.na(Family)) #74 mito

# How to access data in a phyloseq object?
tax_table(raw_physeq)

# How to access data
View(sample_data(raw_physeq))

# Intuition Check
ntaxa(raw_physeq) - ntaxa(noChloroMitos_physeq)
```

#Negative Controls
```{r neg-controls}
#create a vector of samples that were negative controls
control_samples <- c("WaterControl", "022um-Control", "3um-Control", "DNA-Ext-Control")

#create phyloseq object of ONLY negative controls

control_physeq <-
  noChloroMitos_physeq %>%
  subset_samples(names %in% control_samples) %>%
  prune_taxa(taxa_sums(.) > 0, .)

#Visualize the ASV abundances in negative controls
control_physeq %>%
  plot_bar(., "ASV", fill = "ASV")

# Vector of control ASVs (31)

# Create phyoloseq object of just those 31 ASVs across 95 samples
control_ASVs <-
  control_physeq %>%
  taxa_names()

controlASV_physeq<-
  noChloroMitos_physeq %>%
  subset_taxa(ASV %in% Control_ASVs)

controlASV_physeq %>%
  plot_bar(., "ASV", fill = "ASV", facet_grid = Sample_or_Control~.)

# Remove 29 of the control ASVs from the data
controlASVs_toREMOVE <- 
  dplyr::setdiff(control_ASVs, c("ASV_0019", "ASV_0055"))

length(controlASVs_toREMOVE)

# Make a new physeq object without these 29 ASVs
noControlsChlorosMitos_physeq <- 
  noChlorosMitos_physeq %>%
  # actually remove the control ASVs 
  subset_taxa(., !(ASV %in% controlASVs_toREMOVE)) %>%
  # Remove control samples 
  subset_samples(., !(names %in% control_samples))

```


